---
title: "USA Covid Data"
author: "Amritha Girish"
date: "15/08/2021"
output: html_document
---
To load the data set:
```{r}
data=read.csv("USA Covid Data.csv")
```

To display the first 6 rows of the data set
```{r}
head(data)
```
To show the number of specified rows i.e for n=8, it will display the first 8 rows
```{r}
head(data, n=8)

```
To display the last Rows of your data
```{r}
tail(data)
```
To view the previously Loaded named object we use ls() command
```{r}
ls()
```
for examining the data structure of an object str() command can be used
```{r}
str(data)

```
to pick out item from third row and third columns:
```{r}
data[3,3]
```
to display the values of third row corresponding to columns from one to four:
```{r}
data[3,1:4]
```
for displaying all the rows by leaving out the first value and selecting just the first column:
```{r}
data[,1]
```
specifying rows and displaying values for all columns:
```{r}
data[c(1,3,5,7),]
```
specifying several rows and using -4 to display all the columns except the fourth:
```{r}
data[c(1,3,5,7),-4]
```
Since the columns are named we can use the name instead of values:
```{r}
data[c(1,3,5,7),'Population']
```
Giving a single value  in the square brackets, this will display the values in third column along with all the rows:
```{r}
data[3]
```
If you append $ and then the name of the required column you will view that column data:
```{r}
data$Total.Deaths
```
You can also use square brackets to select out parts of the item:
```{r}
#this will give the first four values of the column
data$Total.Deaths[1:4]
```
The summary command  gives us some simple statistics about the numeric data:
```{r}
summary(data)
```

To see the summary of a particular column:
```{r}
summary(data$Total.Cases.1.mil.population)
```
To find the mean of a particular column:
```{r}
mean(data$Total.Deaths)
```
To find the median of a particular column:
```{r}
median(data$Total.Deaths)
```
To find variance of a  column:
```{r}
var(data$Total.Deaths)
```
To find standard deviation:
```{r}
sd(data$Total.Deaths)
```
To calculate the Standard deviation but after revoming the NA items with an additional instruction

```{r}
sd(data$Total.Deaths,na.rm=TRUE)
```

```{r}
quantile(data$Total.Deaths)
```
To find 20% quantile:

```{r}
quantile(data$Total.Cases,0.2)
```
To find 20%,50% and 80% quantile:
```{r}
quantile(data$Total.Cases,c(0.2,0.5,0.8))
```


To Returns five number summary (minimum, lower-hinge, median, upper-hinge, maximum) for the input data.
```{r}
fivenum(data$Total.Cases)
```


The most basic command that enables the viewing of columns or row is name():
```{r}
names(data)
```
For displaying the row names:
```{r}
row.names(data)
```
For displaying the column names:
```{r}
colnames(data)
```
To display both row names and the column names at the same time we can use the command dimnames() 
```{r}
dimnames(data)
```
Omitting NA items:
```{r}
length(na.omit(data$Total.Recovered))
```
Interpreting the type of object using the class() command:

```{r}
class(data)
```
```{r}
class(data$Total.Deaths)
```
```{r}
class(dimnames(data))
```
Rotating the data frames, rows become columns and vice versa.
```{r}
t(data)
```
The cbind() and rbind() commands assemble a matrix/data frames, by columns or rows, from several other object

```{r}
head(rbind(data))

```
```{r}
head(cbind(data))
```

Selecting and displaying parts of contingency table:
```{r}
data.tab=with(data,table(Total.Cases,Active.Cases))
data.tab
```
Now to display only 3 rows of the contingency table:
```{r}
data.tab[1:3,]
```
Examine the table object using str() command:

```{r}
str(data.tab)
```
To determine the number of items in an object use length() command:
```{r}
length(data)
```
The max() and min() commands display the largest and smallest values in a numeric object.
```{r}
max(data$Total.Cases)
```
```{r}
row.names(data)[which(data$Total.Cases==3983692)]
#California has the highest reported covid cases

```

```{r}
min(data$Total.Cases)
```
```{r}
row.names(data)[which(data$Total.Cases==25038)]
#Vermont has the lowest reported covid cases

```
```{r}
max(data$Total.Deaths)

```
```{r}
row.names(data)[which(data$Total.Deaths==64543)]
#california has highest deaths reported.
```
```{r}
min(data$Total.Deaths)
```
```{r}
row.names(data)[which(data$Total.Deaths==260)]
#Vermont has the lowest reported deaths.
```

We can sort the values in ascending order using command sort():
```{r}
head(sort(data$Total.Recovered))
```
Converting an object into a table:
```{r}
d.tab=as.table(as.matrix(data))
d.tab
```
Testing for table object:
```{r}
is.table(data)
```

```{r}
is.table(d.tab)
```
Cumulative Statistics:
```{r}
head(cumsum(data$Total.Cases))
```
```{r}
head(cummax(data$Total.Deaths))
```
```{r}
head(cummin(data$Total.Deaths))
```
```{r}
head(cumprod(data$Total.Deaths))
```
Testing whether the data extracted is data frame or not:
```{r}
is.data.frame(data)
```
The if command is used in logical testing of some conditions and can be paired with else to provide an alternative:
```{r}
if(class(data) == 'data.frame') TRUE else FALSE
```
```{r}
if(class(d.tab) == 'data.frame') TRUE else FALSE
```
```{r}
if(class(data.tab) == 'table') TRUE else FALSE
```
We can use seq_along() command to create a simple index:
```{r}
seq_along(data)
```

Looking at the distribution of the data:

stem and leaf plot: FRequency is represented in right and range of numeric categories on left.
```{r}
stem(data$Total.Deaths)
```
Histograms can be created using graphical command hist():
```{r}
hist(data$Total.Deaths)
```
```{r}
hist(data$Total.Deaths,col='purple3',main=NULL,xlab='total deaths')
```
```{r}
hist(data$Total.Cases,col='yellow3')
```
```{r}
hist(data$Population)
```
Density function:
Instead of using frequency=FALSE in the hist() command we can directly use density() command.
```{r}
density(data$Total.Cases)
```
Using density function to Draw a graph:
```{r}
plot(density(data$Total.Cases))
```
Adding density lines to the Existing graphs:
```{r}
hist(data$Total.Cases, freq = F)
lines(density(data$Total.Cases), lty = 2)
lines(density(data$Total.Cases, k = 'rectangular'))
```
Commands you can use in relation to the normal distribution:
```{r}
#to generate n random numbers from the normal distribution with mean of 0 and standard deviation of 1.

head(rnorm(data$Total.Cases, mean = 0, sd = 1))
```
```{r}
# for Returning the probability for the quantile q.
head(pnorm(data$Total.Cases, mean = 0, sd = 1))
```
By using rnorm() we can create ideal normal distribution using the mean and standard deviation of the data.
```{r}
data.norm=rnorm(1000,mean(data$Total.Cases,sd(data$Total.Cases)))
hist(data.norm, freq = FALSE)
lines(density(data.norm))

```
Quantile-Quantile Plot: 
It produces a series of points that appers in a straight line for normal distribution. 
```{r}
qqnorm(data$Total.Cases)# for non-normal distribution

```
```{r}
qqnorm(data.norm)# for normal distribution
```
```{r}
data.norm2=rnorm(1000,mean(data$Total.Deaths,sd(data$Total.Deaths)))
qqnorm(data.norm2)
```
```{r}
qqnorm(data.norm, main = 'QQ plot of data$total.cases', xlab = 'Theoretical', 
ylab = 'Quantiles for data$total.cases')
qqline(data.norm, lwd = 2, lty = 2)
```
It is possible to plot one distribution against another as a quantile-quantile plot using the qqplot() command.
```{r}
qqplot(data$Total.Cases, data$Total.Deaths)

```
```{r}
qqplot(data.norm, data.norm2)
```
Drawing a straight line on the qqplot() can be done using the abline() command. This command uses the properties of a straight line (that is, y = a + bx) to produce a line on an existing plot. The general form of the command is: abline(a = intercept, b = slope).
```{r}
qqp = qqplot(data$Total.Cases, data$Total.Deaths)
abline(lm(qqp$y ~ qqp$x))
```
```{r}
#for normalized data
qqp = qqplot(data.norm, data.norm2)
abline(lm(qqp$y ~ qqp$x))
```
The basic method of applying a t-test is to compare two vectors of numeric data

```{r}
t.test(data$Total.Cases, data$Total.Deaths)
```
Two-sample t-Test with unequal variance:
```{r}
t.test(data$Total.Cases,data$Total.Deaths)
```
Two-sample t-Test with Equal variances:
This can override the default and use the classic t-test by adding the var.equal = TRUE instruction, which forces the command to assume that the variance of the two samples is equal.
```{r}
t.test(data$Total.Cases, data$Total.Deaths, var.equal = TRUE)
```
One sample t-Testing:
In this version you supply the name of a single vector and the mean to compare it to (this defaults to 0):
```{r}
t.test(data$Total.Cases, mu = 5)
```
Using Directional Hypotheses:
In this the direction of the hypotheses can be specified.
```{r}
t.test(data$Total.Cases, mu = 5, alternative = 'greater')

```
Two-Sample U-Test:
```{r}
wilcox.test(data$Total.Cases, data$Total.Deaths)
```
One-Sample U-Test:
```{r}
wilcox.test(data$Total.Cases,exact=FALSE)
```
Correlation and Covariance:
The cor() command determines correlation and cov()command determines covariance.
```{r}
#correlation
cor(data$Total.Cases, data$Total.Deaths)

```
```{r}
#Determines covariance between x and y . If x is a matrix or data frame, y can be omitted.
x=cov(data$Total.Cases, data$Total.Deaths)
x

```
The cov2cor() command is used to determine the correlation from a matrix of covariance in the following example:
```{r}
v = as.matrix(x)
cov2cor(v)
```
Significance Testing in Correlation Tests:
```{r}
cor.test(data$Total.Cases, data$Total.Deaths)
```
Chi-squared tests of association can be carried out using the chisq.test() command.
```{r}
chisq.test(data$Total.Cases)

```
BoxPlot’s:
```{r}
boxplot(data$Total.Cases)
```
```{r}
boxplot(data$Total.Deaths)
```
```{r}
boxplot(data$Total.Recovered)

```

Boxplot for multiple items:
```{r}
boxplot(data$Total.Cases, data$Total.Deaths)
```
```{r}
boxplot(data$Total.Cases~data$Total.Deaths , data = data)
```
Customising Boxplots:
```{r}
boxplot(data$Total.Cases,data$Active.Cases,names=c('Total cases','Active cases'))
title(xlab="Variable",y="Value")
```
Horizontal Boxplots:
```{r}
boxplot(data$Total.Cases, horizontal = T)
title(xlab = 'Total caes')
```
```{r}
boxplot(data$Total.Deaths, horizontal = T)
title(xlab = 'Total deaths')
```
```{r}
boxplot(data$Total.Recovered, horizontal = T)
title(xlab = 'Total Recovered')
```
scatter plot’s:
```{r}
library(dplyr)
```
```{r}
plot(data$Total.Cases,data$Active.Cases)
```
```{r}
plot(data$Total.Recovered,data$Active.Cases)

```
Pairs plots(multiple correlation plots)
```{r}
plot(data)
```
```{r}
a=head(data)
plot(a)
```
 Desired columns can be selected for ploating:
```{r}
 pairs(~ Total.Cases + Active.Cases + Total.Recovered, data =data)
```
Line Charts:
```{r}
 with(data[order(data$Total.Cases),], plot(sort(Total.Cases), Active.Cases, type = 'l'))
```
```{r}
plot(data$Total.Cases, type = 'l' )
```
```{r}
plot(data$Total.Cases, type = 'b')
```
Pie chartS:
```{r}
pie(data$Total.Cases, labels = 'Total cases')
```
```{r}
pie(data$Total.Deaths, labels = 'Total deaths')
```
Cleveland Dot Charts:
```{r}
dotchart(data$Total.Cases)
```
```{r}
dotchart(data$Total.Deaths)
```
```{r}
dotchart(data$Active.Cases)
```
```{r}
dotchart(data$Total.Recovered)
```
```{r}
data.mat=head(as.matrix(data$Total.Deaths))
dotchart(head(data$Total.Deaths))
```
Use the t() command to transpose the matrix and produce your dot plot:
```{r}
dotchart(t(data.mat))
```
Bar charts:
The bar chart is suitable for showing data that fall into discrete categories.
```{r}
barplot(data$Total.Cases)
```
```{r}
barplot(data$Total.Cases)
title(xlab = 'states', ylab = 'total cases')
box()
```
```{r}
barplot(data$Total.Cases, names = row.names(data), ylab = 'Count' ,
col = 'tan')
```
Multiple Category Bar Charts:
```{r}
#Stacked Bar Charts
data.na=na.omit(data)
barplot(as.matrix(data.na))
```
Horizontal Bars:
```{r}
barplot(data.mat, beside = TRUE, horiz = TRUE)
```
The analysis of variance using the aov() command
```{r}
temp.aov = na.omit(aov(data$Total.Cases ~ data$Total.Deaths, data = data))
temp.aov
```
To see the classic ANOVA table of results, use the summary() command:
```{r}
summary(temp.aov)
```
Two-Way ANOVA:
In a basic one-way ANOVA you have one response variable and one predictor variable, but you may come across a situation in which you have more than one predictor variable, in such cases use two-way ANOVA.
```{r}
data.aov = aov(Total.Cases ~ Total.Deaths + Population, data = data)
summary(data.aov)
```
creating new testCol/testRow for addition as a new column/row to the df object (using rep() command):
```{r}
testCol = c(rep(data$Total.Cases - data$Total.Deaths, length(data)))
head(testCol)
```
```{r}
testRow = c('Albama1', 2100, 3100, 263, 2888, 18,66, 334, 768, 3000)
testRow
```
Adding Rows or Columns:
```{r}
test = data.frame(data, testCol)
head(test)
```
```{r}
test = data
test['testRow',] = testRow
tail(test)
```
```{r}
test = data
test["Test",] = NA
tail(test)
```
Column Indexes:
```{r}
col(data, as.factor = F)
```
Extracting two columns from data:
```{r}
x = data[ , 9:10]
head(x)
```
Use the apply() command to apply a function over all the rows or columns of a data frame (or matrix):
```{r}
#similar to colMean()
apply(data[ , 9:10], 2, mean)
```
```{r}
apply(data[ , 5:10], 2, mean)
```
```{r}
apply(data[ , 9:10], 2, median)
```
```{r}
apply(data[ , 9:10], 2, var)
```
Using tapply() to Summarize Using a Grouping Variable:
```{r}
tapply(data$Total.Cases, data$State, FUN = mean)

```
```{r}
tapply(data$Total.Deaths, data$State, FUN = mean)
```
```{r}
tapply(data$Total.Recovered, data$State, FUN = mean)
```
```{r}
tapply(data$Active.Cases, data$State, FUN = mean)
```
```{r}
tapply(data$Population, data$State, FUN = mean)
```
```{r}
tapply(data$Total.Tests, data$State, FUN = mean)
```
The aggregate() command enables you to compute summary statistics for subsets of a data frame or matrix.
```{r}
aggregate(data[ , 9:10], by = list(data$State), FUN = mean)
```
```{r}
aggregate(data[ , 2:3], by = list(data$State), FUN = mean)
```
```{r}
aggregate(data[ , 3:4], by = list(data$State), FUN = mean)
```
```{r}
aggregate(data[ , 3:6], by = list(data$State), FUN = mean)
```
Using cbind():
```{r}
aggregate(cbind(data$Total.Cases, data$Total.Recovered), data = data, by = list(data$State), FUN = mean)
```
```{r}
aggregate(data$Total.Cases ~ data$State* data$Total.Recovered, data = data, FUN = mean)
```
```{r}
aggregate(data$Total.Deaths ~ data$Total.Cases, data = data, FUN = mean)
```
objects returns a vector of character strings giving the names of the objects in the specified environment:
```{r}
objects(data)
```
ls() is similar to object():
```{r}
ls(data)
```
Simple lineaR RegreSSion:
```{r}
datalr = lm(Total.Deaths ~ State, data = data)
datalr

```
```{r}
cor.test(~ data$Total.Deaths + data$Total.Cases, data = data)
```
Linear Model Results Objects: coefficients can be extracted using the coef() command.
```{r}
coef(datalr)
```
Fitted Values
```{r}
head(fitted(datalr))
```
Formula
```{r}
formula(datalr)
```
```{r}
datalr$call
```
```{r}
formula(data)
```
Using the segments() Command for Error Bars:
```{r}
data.m = apply(data[,9:10], 2, mean)
data.m
```
```{r}
data.sum = apply(data[,9:10], 2, sum)
data.sum
```
```{r}
data.sd = apply(data[ , 9:10], 2, sd)
data.sd
```
```{r}
data.1 = data.sum/data.m
data.1
```
```{r}
data.se = data.sd / sqrt(data.1)
data.se
```
```{r}
data.m + data.se
```
```{r}
max(data.m + data.se)
```
```{r}
data.max = round(max(data.m + data.se) + 0.5, 0)
data.max
```

```{r}
bplt = barplot(data.m, ylim = c(0, data.max))
```
Logarithmic Regression:
```{r}
 plot(head(Total.Cases) ~ head(Total.Recovered), data = data)

```
Polynomial Regression:
```{r}
 plot(Total.Cases ~ Total.Recovered, data = data)
```
```{r}
data.lm = lm( Total.Cases ~ Total.Recovered + I(Total.Recovered^2), data = data)
summary(data.lm)
```
Diagnostic Plots:
```{r}
plot(data.lm)
```
Creating Mathematical Expressions:
```{r}
plot(1:10, 1:10, type = 'n')
opt = par(cex = 1.5)
text(1, 1, expression(hat(x)))
text(2, 2, expression(alpha==x))
text(3, 3, expression(beta==y))
text(4, 4, expression(frac(x, y)))
text(5, 5, expression(sum(x)))
text(6, 6, expression(sum(x^2)))
text(7, 7, expression(bar(x) == sum(frac(x[i], n), i==1, n)))
text(8, 8, expression(sqrt(x)))
text(9, 9, expression(sqrt(x, 3)))

```






